TacticToe: Learning the Game of Tactics

### 2018-02-16 tactictoe_jv.tex
commentary on the experiments

%Mention the conflicts.
%Theorem orthogonalization is not turn on during instantiation of tacitcs
%except
%for the metis tactic.
%Best parameters + comparison with holyhammer.

The best \tactictoe parameters are chosen for the full-scale evaluation. Yet,
\eprover asynchronous calls are not included in the best \tactictoe strategy
for fairer comparison.

@@

Table~\ref{theories} compares the success rates of re-proving for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics.

@@

@@

\subsection{Complexity of the proofs}

@@

Estimate the difficulty of the proof by the number of tactic units.

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=south east, at={(0.9,0.1)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=60,
  ymin=0, ymax=4800,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=time, y=solved] {data/tactictoe_time};
\addplot table[x=time, y=solved] {data/eprover_time};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{Number of problem solved in less than x seconds.}
\end{figure}
\eprover uses at least two strategies where as \tactictoe only uses one set
of parameters.

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the proofs are short. The proofs found by tactictoe follows
about the same
distribution. This means that if \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.
In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
see how well \tactictoe and \eprover performs on increased level of
difficulties. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem by their own mean.
The performance of \tactictoe is compelling with more than half of the theorem
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.

@@

>>>>

Based on the results of parameter tuning, we now evaluate a version of 
\tactictoe with its best parameters on an evaluation set of 7164 theorems from 
the \holfour standard library. We compare it with the performance of \eprover.
For reasons of fairness, \eprover asynchronous calls are not included in the 
best \tactictoe strategy. The ATP \eprover is run in auto-schedule mode with 
128 premises. The settings for \tactictoe are the following:
0.05 seconds tactic timeout, an orthogonalization radius of 20, theorem list 
abstraction with 16 predicted theorems for instantiation, a prior policy 
coefficient of 0.5, an evaluation radius of 10, an exploration coefficient of 2 
and priority calls to \metis with a timeout of 0.1 seconds.

@@

The results shows that \tactictoe is able to prove almost twice as much 
theorems as \eprover. Combining the results of \tactictoe and \eprover we get a 
69.0\% success rate which is significantly above the 5O\% success rates 
of hammers on this type of problems~\cite{tgck-cpp15}. Moreover, \tactictoe is 
running a single set of parameters (strategy), where as hammers and ATPs have 
been optimized and rely on a wide range of strategies.

\paragraph{Reconstruction}
Tactic proofs produced by \tactictoe during this experiment are all verifiable 
in \holfour. By the design of the proof search, reconstruction of \tactictoe 
proof succeeds, unless one of the tactic modifies the state of \holfour in a 
way that changes the behavior of a tactic used in the final proof. Such an 
event has not yet occurred.
More details on how a proof is extracted from the search tree is given in 
Section~\ref{sec:proofdisplay}.
For \eprover, reconstruction rate varies between 95 and 99 percent, depending 
on the reconstruction method \cite{}\todo{CK}.

@@

The Table~\ref{theories} compares the success rates of re-proving for different
\holfour theories. \tactictoe outperforms \eprover on every
considered theory.
\eprover is more suited to deal with dense theories such as
\texttt{real} or \texttt{complex} where a lot of related theorems are available
and most proofs are usually completed by rewriting tactics. Thanks to its 
ability to re-use custom-built tactics, \tactictoe 
largely surpasses \eprover on theories relying on inductive terms and 
simplifications sets such as \texttt{arithmetic}, \texttt{list} 
and \texttt{f\_map}. Indeed, \tactictoe is able to recognize where and when to 
apply induction, a task at which ATPs are known to struggle with.

@@

\subsection{Complexity of the proof search}

@@

\begin{figure}[h]
\centering
\begin{tikzpicture}[scale=1]
\begin{axis}[
  legend style={anchor=south east, at={(0.9,0.1)}},
  width=\textwidth,
  height=0.7*\textwidth,xmin=0, xmax=60,
  ymin=0, ymax=4800,
  xtick={},
  ytick={},
  cycle list name=my black]
\addplot table[x=time, y=solved] {data/tactictoe_time};
\addplot table[x=time, y=solved] {data/eprover_time};
\legend{\tactictoe,\eprover}
\end{axis}
\end{tikzpicture}
\caption{\label{fig:c1} Number of problems solved in less than x seconds.}
\end{figure}

We first investigate how \tactictoe and \eprover scales as the search time 
grows in Figure~\ref{fig:c1}.
In 10 seconds, \tactictoe solves 90 percent of the problem it can 
solve in 60 seconds. The analysis is a bit different for \eprover. Indeed,
we can clearly deduce from the bump at 30 seconds that \eprover is using at 
least two strategies and probably more. Strategies are a useful method to 
fight the exponential decline in the number of newly proven theorems over 
time. Therefore, integrating strategy scheduling in \tactictoe could be 
something to be experimented with. Overall, we recommend using \tactictoe and 
\eprover with a 10 seconds timeout. It is possible also to run them with a 
longer time limit, but we advise to do so in parallel to avoid waiting time.

@@

In order to appreciate the difficulty of the evaluation set from a human
perspective, the length distribution of human proofs in the
validation set is shown in Figure~\ref{fig:distrib}.  It is clear from the
graph that most of the human proofs are short. The proofs found by tactictoe 
follows about the same
distribution. This means that if \tactictoe finds a proof, there is about a 50
percent chance that it will be shorter than what a human would come up with.

@@

In Figure~\ref{fig:percentage}, we regroup the human proofs by their length to
see how well \tactictoe and \eprover performs on increased level of
difficulties. As expected, the longer the original proof is, the harder it is
for \tactictoe and \eprover to re-prove the theorem by their own mean.
The performance of \tactictoe is compelling with more than half of the theorem
that required a proof of length six being reproven. It is also consistently
better than \eprover for any proof length.
